---
title: "Phyloseq"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load required packages
```{r}
library(phyloseq)
library(dplyr)
library(BiMiCo)
library(ggplot2)
```
#load taxa and seqtab.nochim
```{r}
load("RData/taxa.RData")
load("RData/seqtab.nochim.RData")
```
# import metadata
```{r}
metadata<-read.csv("metadata.csv", header=TRUE, row.names = 1)
```
#create phyloseq object
```{r}
phyloseq<- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), sample_data(metadata), tax_table(taxa))
physeq
```
# remove the sequence itself and replace with ASV
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(physeq))
names(dan) <- taxa_names(physeq)
physeq <- merge_phyloseq(physeq, dna)
taxa_names(physeq) <- paste0("ASV", seq(ntaxa(physeq)))
physeq
```
#remove the mitochondria and chloroplast matches.  remove all non bacterial sequences
```{r}
physeq<- physique %>% subset_taxa(Family!= "Mitochondria" | is.na(Family) & Order!="Chloroplast" | is.na(Order))
physeq

```
# remove all non bacterial sequences
```{r}
physeq<-rm_nonbac(physeq)
physeq
```
#save physeq objects to load later
```{r}
save(physeq, file="RData/physeq.RData")
```
#load physeq objects to start here
```{r}
load("RData/physeq.RData")
```
# plot bar graph based on phylum 
```{r}
plot_bar(physeq, fill="Phylum") + geom_bar(aes(color=Phylum, fill=Phylum), stat="identity", position="fill")
```
# create a barplot of relative abundance
```{r}
physeq_relabund <- transform_sample_counts(physeq, function(x) x / sum(x))
```
# barplot
```{r}
plot_bar(physeq_relabund, fill="Phylum") + geom_bar(aes(color=Phylum, fill=Phylum), stat="identity", position="fill") + facet_wrap(~body.site, scales= "free")
```

# plot alpha diversity based on body.site
```{r}
plot_richness(physeq, x="body.site", color="subject", measures=c("Observed", "Simpson", "Shannon"))
```
#plot alpha diversity based on reported.antibiotic.usage
```{r}
plot_richness(physeq, x="reported.antibiotic.usage", color="body.site", measures=c("Observed", "Simpson", "Shannon"))
```
#plot alpha diversity based on subject
```{r}
plot_richness(physeq, x="subject", color="body.site", measures=c("Observed", "Simpson", "Shannon"))
```
# test for normality
```{r}
alpha <- estimate_richness(physeq, measures=c("Observed", "Simpson", "Shannon"))
```
# shapiro-wilk
```{r}
observed <- shapiro.test(alpha$Observed)
shannon <- shapiro.test(alpha$Shannon)
simpson <- shapiro.test(alpha$Simpson)
```
#print
```{r}
print(observed)
print(shannon)
print(simpson)
```

#create data frames for statistical analyses
```{r}
samples <- sample_data(physeq)
if (class(samples) =="sample_data") {samples <- data.frame(sample_data(samples))}
alpha$sample <- rownames(alpha)
alpha <- merge(alpha, samples, by = "sample")
```
# perform statistics based on subject
```{r}
test_observed <- wilcox.test(Observed ~ subject, data=alpha)
test_simpson <- wilcox.test(Simpson ~ subject, data=alpha)
test_shannon <- t.test(Shannon ~ subject, data=alpha)
print(test_observed)
print(test_simpson)
print(test_shannon)
```

# perform statistics based on reported.antibiotic.usage
```{r}
test_observed <- wilcox.test(Observed ~ reported.antibiotic.usage, data=alpha)
test_simpson <- wilcox.test(Simpson ~ reported.antibiotic.usage, data=alpha)
test_shannon <- t.test(Shannon ~ reported.antibiotic.usage, data=alpha)
print(test_observed)
print(test_simpson)
print(test_shannon)
```
# test for body site
```{r}
kruskal.test(Simpson ~ body.site, data=alpha)
pairwise.wilcox.test(alpha$Simpson, alpha$body.site, p.adjust.method = "holm")
kruskal.test(Observed ~ body.site, data=alpha)
pairwise.wilcox.test(alpha$Observed, alpha$body.site, p.adjust.method = "holm")
shannonanova <- aov(Shannon ~ body.site, data=alpha)
summary(shannonanova)
TukeyHSD(shannonanova)
```
# remove taxa with relative abundance <0.005%
```{r}
minTotRelAbun = .00005
x = taxa_sums(physeq)
keepTaxa = (x/sum(x)) > minTotRelAbun
Phyloseqprune = prune_taxa(keepTaxa , physeq)
Physeqprune
```
# NUMBER OF SHARED ASVS BODY.SITE (Found in 25% or more)
```{r}
bodysite=ps_venn(
Physeqprune,
"body.site" ,
fraction = .25,
weight = FALSE,
relative = TRUE,
Plot = TRUE
)
bodysite
```
# Bray Curtis calculation
```{r message =FALSE}
set.seed(777)
dist = phyloseq::distance(physeqprune, method= "bray", weighted = TRUE)
ordination = ordinate(physeqprune, method = "NMDS", distance=dist)
```
# Bray Curtis Subject Plot
```{r}
braysubject=plot_ordination(physeqprune, ordination, color="subject") +
theme_classic() + 
theme(strip.background = element_blank()) + stat_ellipse(aes(group=subject))
braysubject
```
# Bray Curtis Body site Plot
```{r}
braybodysite=plot_ordination(physeqprune, ordination, color = "body.site") + 
theme_classic() +
theme(strip.background = element_blank()) + stat_ellipse(aes(group=subject))
braysubject
```
# Bray Curtis reported antibiotic usage  (fix code  for this section)
```{r}
braybodysite=plot_ordination(physeqprune, ordination, color = "body.site") + 
theme_classic() +
theme(strip.background = element_blank()) + stat_ellipse(aes(group=subject))
braysubject
```
#Bray Curtis Subject Statisitcs
```{r}
adonisZ(dist - sample_data(physeqprune)$subject)
```
# Bray Curtis site statistics
```{r}
adonisZ(dist - sample_data(physeqprune)$body.site)
Ps.disper<-betadisper(dist, sample_data(physeqprune)$body.site)
permutest(ps.disper, pair=TRUE)
```
